---
title: "Case_2"
author: "Alexandre Martins Fontes"
date: "04/07/2020"
output: pdf_document
---


**1)Show the evolution of total monthly and total annual exports from Brazil (all states and to everywhere) of ‘soybeans’, ‘soybean oil’ and ‘soybean meal’**

Imports were also included in plots for a supperficial comparation between magnetudes.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)

#####CODE WITH Ggplot2#####
rm(list = ls()) 

library(forecast)
library(ggplot2)
library(readxl)
library(hrbrthemes)
library(dplyr)


data= read.csv('data_comexstat.csv')
covariates = read_excel('covariates.xlsx')



#soybeans by year

data2 = subset(data, data$product=='soybeans' )

data2$date = as.Date(data2$date)

p <- ggplot(data2, aes(x = data2$date, y = data2$usd , color = as.factor(data2$type)) ) +
  geom_point(size = 0.5) +
  scale_x_date(date_breaks = "years", date_labels = "%Y")+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Soybeans imports and exports", x = "Data (years)", y = "Value (US$)", color = "Type of trade")

p

```

## Data manipulation for Soybeans by year plot

All the data manipulation are avaliable and also can be explored/reproduced in the file case_2.R


```{r}
knitr::opts_chunk$set(echo = T)

#####CODE WITH DATA MANIPULATION#####
#soybean-meal by year
rm(list = 'data2') 

data2 = subset(data, data$product=='soybean_meal' )

data2$date = as.Date(data2$date)

p <- ggplot(data2, aes(x = data2$date, y = data2$usd , color = as.factor(data2$type)) ) +
  geom_point() +
  scale_x_date(date_breaks = "years", date_labels = "%Y")+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Soybean meal imports and exports", x = "Data (years)", y = "Value (US$)", color = "Type of trade")

p

```
## Data manipulation for Soybean meal by year plot

All the data manipulation are avaliable and also can be explored/reproduced in the file case_2.R



```{r}
knitr::opts_chunk$set(echo = T)

#####CODE WITH DATA MANIPULATION#####
#soybean-oil by year
rm(list = 'data2') 

data2 = subset(data, data$product=='soybean_oil' )

data2$date = as.Date(data2$date)

p <- ggplot(data2, aes(x = data2$date, y = data2$usd , color = as.factor(data2$type)) ) +
  geom_point(size = 0.5) +
  scale_x_date(date_breaks = "years", date_labels = "%Y")+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Soybean oil imports and exports", x = "Data (years)", y = "Value (US$)", color = "Type of trade")

p
```
## Data manipulation for Soybean oil by year plot

All the data manipulation are avaliable and also can be explored/reproduced in the file case_2.R


```{r}
knitr::opts_chunk$set(echo = T)

#####CODE WITH DATA MANIPULATION#####
#soybeans by month
rm(list = 'data2') 

data2 = subset(data, data$product=='soybeans' )

data2$date = as.numeric(substr ( data2$date, 6,7 ))




p <- ggplot(data2, aes(x = data2$date, y = data2$usd , color = as.factor(data2$type)) ) +
  geom_point(size = 0.5) +
  scale_x_continuous(breaks = seq(1,12, by = 1))+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Soybeans imports and exports by month", x = "Months", y = "Value (US$)", color = "Type of trade")

p

```
## Data manipulation for soybeans  by month plot

All the data manipulation are avaliable and also can be explored/reproduced in the file case_2.R

```{r}
knitr::opts_chunk$set(echo = T)

#####CODE WITH DATA MANIPULATION#####
#soybean meal by month
rm(list = 'data2') 

data2 = subset(data, data$product=='soybean_meal' )

data2$date = as.numeric(substr ( data2$date, 6,7 ))




p <- ggplot(data2, aes(x = data2$date, y = data2$usd , color = as.factor(data2$type)) ) +
  geom_point(size = 0.5) +
  scale_x_continuous(breaks = seq(1,12, by = 1))+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Soybean meal imports and exports by month", x = "Months", y = "Value (US$)", color = "Type of trade")

p

```
## Data manipulation for soybean meal  by month plot

All the data manipulation are avaliable and also can be explored/reproduced in the file case_2.R


```{r}
knitr::opts_chunk$set(echo = T)

#####CODE WITH DATA MANIPULATION#####
#soybean oil by month
rm(list = 'data2') 

data2 = subset(data, data$product=='soybean_oil' )

data2$date = as.numeric(substr ( data2$date, 6,7 ))




p <- ggplot(data2, aes(x = data2$date, y = data2$usd , color = as.factor(data2$type)) ) +
  geom_point(size = 0.5) +
  scale_x_continuous(breaks = seq(1,12, by = 1))+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Soybean oil imports and exports by month", x = "Months", y = "Value (US$)", color = "Type of trade")

p

```
## Data manipulation for soybean oil  by month plot

All the data manipulation are avaliable and also can be explored/reproduced in the file case_2.R




**2) What are the 3 most important products exported by Brazil in the last 5 years?**

```{r}
knitr::opts_chunk$set(echo = T)
#3 most important products
rm(list = 'data2') 

data2 = subset(data, data$type=='Export' )
data2$date = as.Date(data2$date)
data2= data2[ data2$date > as.Date("2014-12-01"), ]

corn_5 = sum(data2$usd[data2$product=='corn'])
soybean_meal_5 = sum(data2$usd[data2$product=='soybean_meal'])
soybean_oil_5 = sum(data2$usd[data2$product=='soybean_oil'])
soybeans_5 = sum(data2$usd[data2$product=='soybeans'])
sugar_5 = sum(data2$usd[data2$product=='sugar'])
wheat_5 = sum(data2$usd[data2$product=='wheat'])

three_most_important = c(soybeans_5,sugar_5,soybean_meal_5)
p <- ggplot(data2, aes(x = data2$date, y = data2$usd , color = as.factor(data2$product)) ) +
  geom_point(size = 1) +
  scale_x_date(date_breaks = "years", date_labels = "%Y")+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Exports by product in the last 5 years", x = "Data (years)", y = "Value (US$)", color = "Product")

p
```


It is clear from the graph that 'soybeans' is the product most responsible for financial movements in the last 5 years.

The second and third most important products are 'sugar' and 'soybean meal'



**3) What are the main routes through which Brazil have been exporting ‘corn’ in the last few years? Are there differences in the relative importancem of routes depending on the product?**

```{r}
knitr::opts_chunk$set(echo = T)

rm(list = 'data2','soybeans_5','sugar_5','soybean_meal_5','soybean_oil_5','corn_5') 

data2 = subset(data, data$type=='Export' )
data2$date = as.Date(data2$date)
data2= data2[ data2$date > as.Date("2014-12-01"), ]
cor = table(data2$route[data2$product=='corn'])
pie(table(data2$route[data2$product=='corn']), main = "Corn routes", col = rainbow(length(cor)))


```


Corn is exported in the last years mostly by sea


```{r}
knitr::opts_chunk$set(echo = T)

pie(table(data2$route[data2$product=='soybeans']), main = "Soybeans routes", col = rainbow(length(cor)))
pie(table(data2$route[data2$product=='soybean_meal']), main = "Soybean meal routes", col = rainbow(length(cor)))
pie(table(data2$route[data2$product=='soybean_oil']), main = "Soybean oil routes", col = rainbow(length(cor)))
pie(table(data2$route[data2$product=='sugar']), main = "Sugar routes", col = rainbow(length(cor)))
pie(table(data2$route[data2$product=='wheat']), main = "Wheat routes", col = rainbow(length(cor)))



```
With the exception of soybean oil, all products have their routes mostly by sea.


**4)Which countries have been the most important trade partners for Brazil in terms of ‘corn’ and ‘sugar’ in the last 3 years?**

```{r}
rm(list = 'data2')
data2 = subset(data, data$product=='corn' )
data2$date = as.Date(data2$date)
data2= data2[ data2$date > as.Date("2017-12-01"), ]
temp = as.data.frame(table(data2$country)>50)
temp = subset(temp,temp==TRUE)

print(temp)
```

Here we have a list of the trade partners that had at least 50 transations in terms of "corn" in the last 3 years

```{r}
rm(list = 'data2')
data2 = subset(data, data$product=='sugar' )
data2$date = as.Date(data2$date)
data2= data2[ data2$date > as.Date("2017-12-01"), ]
temp = as.data.frame(table(data2$country)>100)
temp = subset(temp,temp==TRUE)

print(temp)
```

Here we have a list of the trade partners that had at least 100 transations in terms of "sugar" in the last 3 years



**5)For each of the products in the dataset, show the 5 most important states in terms of exports?**

```{r}
rm(list = 'data2')
data2 = subset(data, data$type=='Export' )
sugar_state =sort( table(data2$state[data2$product=='sugar']),decreasing = TRUE)
print(sugar_state[1:5])
wheat_state =sort( table(data2$state[data2$product=='wheat']),decreasing = TRUE)
print(wheat_state[1:5])
soybeans_state =sort( table(data2$state[data2$product=='soybeans']),decreasing = TRUE)
print(soybeans_state[1:5])
soybean_oil_state =sort( table(data2$state[data2$product=='soybean_oil']),decreasing = TRUE)
print(soybean_oil_state[1:5])
soybean_meal_state =sort( table(data2$state[data2$product=='soybean_meal']),decreasing = TRUE)
print(soybean_meal_state[1:5])
```


**6)Now, we ask you to show your modelling skills. Feel free to use any type of modelling approach, but bear in mind that the modelling approach depends on the nature of your data, and so different models yield different estimates and forecasts. To help you out in this task we also provide you with a dataset of possible covariates (.xlsx). They all come from public sources (IMF, World Bank) and are presented in index number format. Question: What should be the total brazilian soybeans, soybean_meal, and corn export forecasts, in tons, for the next 11 years (2020-2030)? We’re mostly interested in the annual forecast.**


```{r}
rm(list = 'data2', 'temp')
data2 = subset(data, data$type=='Export' )
data2 = subset(data2, data2$product=='soybeans' | data2$product=='soybean_meal' | data2$product=='corn' )
data2$date = as.numeric(substr ( data2$date, 1,4 ))
year = matrix(2020:2030, 10,1)



#soybeans_forecast
temp = cbind(data2$date[data2$product=="soybeans"],data2$tons[data2$product=="soybeans"])
colnames(temp) = c("year","tons")
temp = as.data.frame(temp)
temp = group_by(temp,year)%>%summarise(tons=sum(tons))
temp2=as.data.frame(forecast(temp$tons, h=10))[1]
temp2=cbind(year,temp2)
dataSOYBEANS = as.data.frame(rbind(as.matrix(temp),as.matrix(temp2)))
product = matrix(rep('soybeans'),33,1)
dataSOYBEANS = cbind(dataSOYBEANS,product)
rownames(dataSOYBEANS) <- NULL

#soybean_meal_forecast
temp = cbind(data2$date[data2$product=="soybean_meal"],data2$tons[data2$product=="soybean_meal"])
colnames(temp) = c("year","tons")
temp = as.data.frame(temp)
temp = group_by(temp,year)%>%summarise(tons=sum(tons))
temp2=as.data.frame(forecast(temp$tons, h=10))[1]
temp2=cbind(year,temp2)
dataSOYBEAN_MEAL = as.data.frame(rbind(as.matrix(temp),as.matrix(temp2)))
product = matrix(rep('soybean_meal'),33,1)
dataSOYBEAN_MEAL = cbind(dataSOYBEAN_MEAL,product)
rownames(dataSOYBEAN_MEAL) <- NULL

#corn_forecast
temp = cbind(data2$date[data2$product=="corn"],data2$tons[data2$product=="corn"])
colnames(temp) = c("year","tons")
temp = as.data.frame(temp)
temp = group_by(temp,year)%>%summarise(tons=sum(tons))
temp2=as.data.frame(forecast(temp$tons, h=10))[1]
temp2=cbind(year,temp2)
dataCORN = as.data.frame(rbind(as.matrix(temp),as.matrix(temp2)))
product = matrix(rep('corn'),33,1)
dataCORN = cbind(dataCORN,product)
rownames(dataCORN) <- NULL

#Binding and plotting
dataFinal=as.data.frame(rbind(dataCORN,dataSOYBEAN_MEAL,dataSOYBEANS))

p <- ggplot(dataFinal, aes(x = dataFinal$year, y = dataFinal$tons , color = as.factor(dataFinal$product)) ) +
  geom_point() +
  geom_smooth(method = "loess")+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Forecasts up to 2029 (in tons)", x = "Data (years)", y = "Weight (tons)", color = "Product")

p

```

Projecting the data with the 'forecast' library; in 2029 we can expect to export 1.489081e + 08 ton of Soybeans, 1.653039e+07 ton of Soybean meal and 6.331813e+07 ton of Corn.
